Trans-Dimensional Inference for Gravitational Lens Substructures

### 2015-06-16 Papers/ms.tex
Mention potential correction
A complete abstract

\begin{abstract}
We introduce a Bayesian solution to the problem of inferring the density
profile of strong gravitational lenses when the lens galaxy may contain
multiple dark or faint substructures. The source and lens models are based on
a superposition of an unknown number of non-negative basis functions
(or ``blobs'') which were chosen with speed as a primary criterion.
The prior distribution for the blobs' properties is specified hierarchically,
so the ``mass function'' of substructures is a natural output. We use reversible
jump Markov Chain Monte Carlo (MCMC) within Diffusive Nested Sampling (DNS) to
sample the posterior distribution and evaluate the marginal likelihood of the
model. We demonstrate the method on a simulated data set, and also apply it to
the Cosmic Horseshoe, finding ...
\end{abstract}

@@

\subsection{The Lens}
The surface mass density profile of the lens is modelled as a superposition
of a singular isothermal ellipsoid plus external shear (SIE+$\gamma$) and
$N$ circular lens ``blobs''. The SIE+$\gamma$ is a simple and widely used lens
model with analytically available deflection angles, and is intended to account
for the bulk of the lensing effect due to the smooth spatial distribution
of (visible and dark) matter in the lens galaxy. Unfortunately, along with
simplicity comes a lack of realism, and generalising the smooth lens model
is an important future step. Ultimately, concentric
mixtures of smooth components may be the most useful for realistic inference
of the density profile of the halo.

>>>>

\begin{abstract}
We introduce a Bayesian solution to the problem of inferring the density
profile of strong gravitational lenses when the lens galaxy may contain
multiple dark or faint substructures. The source and lens models are based on
a superposition of an unknown number of non-negative basis functions
(or ``blobs'') which were chosen with speed as a primary criterion.
The prior distribution for the blobs' properties is specified hierarchically,
so the ``mass function'' of substructures is a natural output. We use reversible
jump Markov Chain Monte Carlo (MCMC) within Diffusive Nested Sampling (DNS) to
sample the posterior distribution and evaluate the marginal likelihood of the
model. We demonstrate the method on a simulated data set with a single
substructure, which is recovered well with moderate uncertainties. We also
apply the method to an image of the ``Cosmic Horseshoe'' lens, and find some
hints of potential substructures. However, we caution that such results could
also be attributed to misspecifications in the model (such as the shape of
the smooth lens component or the point spread function),
which are difficult to guard against in full generality.
\end{abstract}

@@

\subsection{The Lens}
The surface mass density profile of the lens is modelled as a superposition
of a singular isothermal ellipsoid plus external shear (SIE+$\gamma$) and
$N$ circular lens ``blobs''. The SIE+$\gamma$ is a simple and widely used lens
model with analytically available deflection angles, and is intended to account
for the bulk of the lensing effect due to the smooth spatial distribution
of (visible and dark) matter in the lens galaxy. Unfortunately, along with
simplicity comes a lack of realism, and generalising the smooth lens model
is an important future step. Ultimately, concentric
mixtures of smooth components may be the most useful for realistic inference
of the density profile of the halo. A common approach in the lensing
community, that is similar in spirit, is a superposition of an elliptical
power law, external shear, and a pixellated potential correction. However, the
direct use of blobs is closer to the scientific question at hand
when investigating postential dark substructures.
